---
title: "Forest Cover"
output: 
  html_document:
    toc: TRUE
    theme: lumen
    number_sections: TRUE
    code_folding: show
---

# Introduction

The goal of this competition is to predict the forest cover type (the predominant kind of tree cover) from seven different types of tree cover. This study area includes four wilderness areas located in the Roosevelt National Forest of northern Colorado. These areas represent forests with minimal human-caused disturbances, so that existing forest cover types are more a result of ecological processes rather than forest management practices. As a result, the model will need to predict one out of the 7 groups for each area. The training set includeds 15120 observations with 56 variables. The test set is a lot larger than the training set, which means there may be a problem with overfitting. The test set has 565892 observations with 55 variables.

```{r, message = FALSE}
library(caret)
library(tidyverse)
library(DataExplorer)
library(beepr)
library(ggthemes)
library(GGally)
```

# Exploratory Data Analysis

## Data Structure
First, I will look at the structure of the data. As mentioned earlier, there are 15120 observations in the training set and 565,892 observations in the test set. There is one ID column, 10 quantitative variables, 44 indicator variables for the wilderness area and the soil type, and in the training set there is the variable for the cover type. 
```{r}
train <- read.csv("train.csv")
test <- read.csv("test.csv")
glimpse(train)
glimpse(test)
```

## Missing Values

After understanding the structure of the data, I will look to see how many missing values there are so I can understand how much work will need to be done for data manipulation and cleaning. As the plots below show, this is a complete dataset and there are no missing values. Since there is no need to cleaning the data and impute or remove missing values, I can visualize the data and then begin fitting a model. 

```{r}
plot_missing(train[,1:30])
plot_missing(train[, 31:56])
plot_missing(test[,1:30])
plot_missing(test[,31:55])
```


## Response Variable Split

I want to look at the split of the response variable for the training set. From the scatterplot below, there are equal number of observationsn for each category, so there is no need to worry about imbalanced classes in the training data. 

```{r}
# Histogram to show how many of each forest cover
ggplot(train, aes(x = as.factor(Cover_Type))) + 
  geom_bar(aes(fill = as.factor(Cover_Type)), color = "black") + 
  labs(x = "Cover Type", y = "Frequency") +
  theme_clean() +
  scale_x_discrete(labels = c("Spruce/Fir", "Lodgepole", "Ponderosa", "Cottonwood/Willow", " Aspen", "Douglas-fir", "Krummholz")) +
  scale_fill_viridis_d() +
  theme(legend.position = "none")

# The number of patches for each type of forest cover is approximately equal for all types of forest cover
```

## Relationship between Explanatory Variables


```{r}
train.eda <- train %>% group_by(Id) %>% pivot_longer(Soil_Type1:Soil_Type40, "SoilType") %>% filter(value == 1) %>% mutate(SoilType = as.numeric(str_extract(SoilType, "[0-9]+"))) %>% select(-value) %>% pivot_longer(Wilderness_Area1:Wilderness_Area4, "WildernessArea") %>% filter(value == 1) %>% mutate(WildernessArea = as.numeric(str_extract(WildernessArea, "[0-9]+"))) %>% select(-value)

soils <- train.eda %>%
  group_by(SoilType, Cover_Type) %>%
  summarize(Number = n()) %>% arrange(SoilType)

# Stack bar chart of soil type
ggplot(soils, aes(x = as.factor(SoilType), y = Number)) + geom_bar(position = "stack", aes(fill = as.factor(Cover_Type)), stat = "identity") + theme(legend.title = element_blank(), axis.title.y = element_blank()) + labs(x= "Soil Type")

# Side by side bar of soil type
ggplot(soils, aes(x = as.factor(SoilType), y = Number)) + geom_bar(position = "dodge", aes(fill = as.factor(Cover_Type)), stat = "identity") + theme(legend.title = element_blank(), axis.title.y = element_blank()) + labs(x= "Soil Type")

areas <- train.eda %>%
  group_by(WildernessArea, Cover_Type) %>%
  summarize(Number = n()) %>%
  arrange(WildernessArea)

# Stacked bar chart of wilderness area by soil type
ggplot(areas, aes(x = as.factor(WildernessArea), y = Number)) + geom_bar(position = "dodge", aes(fill = as.factor(Cover_Type)), stat = "identity") + theme(legend.title = element_blank(), axis.title.y = element_blank()) + labs(x= "Wilderness Area")


```

```{r}
ggplot(train, aes(x=as.factor(Cover_Type), y=Elevation)) + 
  geom_jitter(alpha = .05, color = "steelblue") +
  geom_boxplot(alpha = .5) +
  labs(x = "Cover Type")

ggplot(train, aes(x=as.factor(Cover_Type), y=Aspect)) + 
  geom_jitter(alpha = .05, color = "steelblue") +
  geom_boxplot(alpha = .5) +
  labs(x = "Cover Type")

ggplot(train, aes(x=as.factor(Cover_Type), y=Slope)) + 
  geom_jitter(alpha = .05, color = "steelblue") +
  geom_boxplot(alpha = .5) +
  labs(x = "Cover Type")

quant_vars <- train %>% select(Elevation, Aspect, Slope, Horizontal_Distance_To_Hydrology, Vertical_Distance_To_Hydrology, Hillshade_9am, Hillshade_Noon, Hillshade_3pm, Horizontal_Distance_To_Fire_Points, Horizontal_Distance_To_Roadways)

library(corrplot)
names(quant_vars) <- c("Elevation", "Aspect", "Slope", "H-Hydrology", "V-Hydrology", 
                       "Hill-9", "Hill-Noon", "Hill-3", "H-Fire", "H-Road")

corrplot(cor(quant_vars), method = "color", type = "upper", diag = FALSE, addCoef.col = "black", number.digits = 2)

ggcorr(quant_vars, label = TRUE, label_size = 3, label_round = 2, label_alpha = TRUE, hjust = 00.75, size = 3,layout.exp = 1)

library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)

# Plot
p1 <- ggplot(train, aes(x = Elevation, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p2 <- ggplot(train, aes(x = Aspect, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p3 <- ggplot(train, aes(x = Horizontal_Distance_To_Hydrology, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="", x = "HD.Hydro") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p4 <- ggplot(train, aes(x = Vertical_Distance_To_Hydrology, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="", x = "VD.Hydro") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p5 <- ggplot(train, aes(x = Hillshade_9am, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="", title = "Density Plot of Elevation by Cover Type", x = "VD.Hydro") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p6 <- ggplot(train, aes(x = Slope, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="", title = "Density Plot of Elevation by Cover Type", x = "VD.Hydro") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p7 <- ggplot(train, aes(x = Hillshade_Noon, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="", title = "Density Plot of Elevation by Cover Type", x = "VD.Hydro") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p8 <- ggplot(train, aes(x = Hillshade_3pm, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="", title = "Density Plot of Elevation by Cover Type", x = "VD.Hydro") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p9 <- ggplot(train, aes(x = Horizontal_Distance_To_Fire_Points, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="", title = "Density Plot of Elevation by Cover Type", x = "VD.Hydro") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )

p10 <- ggplot(train, aes(x = Horizontal_Distance_To_Roadways, y = as.factor(Cover_Type), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(option = "C") +
  labs(y="", title = "Density Plot of Elevation by Cover Type", x = "VD.Hydro") +
  scale_y_discrete(labels = c("Spruce/Fir - 1", "Lodgepole - 2", "Ponderosa - 3", "Cottonwood/Willow - 4", " Aspen - 5", "Douglas-fir - 6", "Krummholz - 7")) + 
  theme_ridges() +
    theme(
      legend.position="none"
    )
library(patchwork)
p1 / p2 
p3 / p4
p5 / p6
p7 / p8
p9 / p10
```



```{r}
myControl <- trainControl(method = "repeatedcv",
                          number = 10)

tunegrid <- expand.grid(eta = .25,
                        max_depth = 3,
                        colsample_bytree = .9,
                        subsample = .8,
                        nrounds = 100,
                        min_child_weight = 1,
                        gamma = .075)
xgbTree.model <- train(as.factor(Cover_Type)~.-Id,
                   data = train,
                   method = "xgbTree",
                   tuneGrid = tunegrid,
                   trControl = myControl,
                   metric = "Accuracy",
                   preProc = c("nzv","zv", "center", "scale")
)
xgbTree.model
beep(sound = 8)
preds <- predict(xgbTree.model, test)
xgbTree <- data.frame(Id = test$Id, Cover_Type = preds)
write_csv(xgbTree, "xgbTree-preds.csv")
```

```{r}

grid <- expand.grid("n.trees" = 1000, "interaction.depth" = 4, "shrinkage" = 0.1, "n.minobsinnode" = 20)
# n.trees should be 1000
# Currently it is 70%
gbm.mod <- train(as.factor(Cover_Type)~.-Id,
                 data = train,
                 method = "gbm",
                 tuneGrid = grid,
                 trControl = myControl,
                 metric = "Accuracy",
                 verbose = FALSE,
                 preProcess = c("nzv", "zv", "center", "scale")
                 )
beep(sound = 8)
gbm.preds <- predict(gbm.mod, test)
gbm <- data.frame(Id = test$Id, Cover_Type = gbm.preds)
write_csv(gbm, "gbm-preds-2.csv")
```

```{r}
svmLinear.model <- train(as.factor(Cover_Type) ~ . -Id,
                         data = train,
                         method = "svmLinear",
                         tuneLength = 4,
                         trControl = myControl,
                         metric = "Accuracy",
                         #verbose = FALSE,
                         preProcess = c("zv", "center", "scale")
                         )
svmLinear.model
beep(sound = 2)
```

```{r}
# score of 0.62083
myControl <- trainControl(method = "repeatedcv",
                          number = 3)

grid <- data.frame("C"=seq(0,100,20))

svmRadial.model <- train(as.factor(Cover_Type) ~ . -Id,
                         data=train,
                         method="svmRadial",
                         tunegrid = grid,
                         trControl=myControl,
                         metric="Accuracy",
                         preProcess = c("zv", "center", "scale"))

beep(sound=5)

preds <- predict(svmRadial.model, test)
beep(sound=5)
svmR <- data.frame(Id = test$Id, Cover_Type = preds)
write_csv(svmR, "svmRadial-preds.csv")
```

```{r}
# This random forest is the best model so far
grid <- expand.grid("mtry" = c(50, 52), splitrule = "extratrees", min.node.size = 1)
rf.model <- train(as.factor(Cover_Type) ~ . -Id,
                  data = train, 
                  method = "ranger",
                  trControl = myControl,
                  tuneGrid = grid,
                  metric = "Accuracy",
                  preProcess = c("zv", "center", "scale"))
rf.model

beep(sound = 8)

rf.preds <- predict(rf.model, test)
rf <- data.frame(Id = test$Id, Cover_Type = rf.preds)
write_csv(rf, "rf-preds.csv")

```

```{r}

# This is the grid that was used to find the best
#grid <- expand.grid("mtry" = c(20, 30, 40, 50), "coefReg" = c(.2,.5, .8, 1)) # Best comes from mtry = 50 and coefReg = .8
grid <- expand.grid("mtry" = 50, "coefReg" = .8)
reg.rf <- rf.model <- train(as.factor(Cover_Type) ~ . -Id,
                  data = train, 
                  method = "RRFglobal",
                  trControl = myControl,
                  tuneGrid = grid,
                  metric = "Accuracy",
                  preProcess = c("zv", "center", "scale"))
reg.rf
beep(sound = 8)

reg.preds <- predict(reg.rf, test)
reg.rf <- data.frame(Id = test$Id, Cover_Type = reg.preds)
write_csv(rf, "reg-rf-preds-2.csv")
```


```{r}

# Join the predictions together to a single data frame
all.preds <- gbm %>% left_join(rf, by = "Id") %>% left_join(reg.rf, by = "Id")

# Set the seed so the random stays the same
set.seed(4171996)

# Pull the observations where at least two agree
all.preds.attempt <- all.preds %>% filter(Cover_Type.x == Cover_Type.y | Cover_Type.y == Cover_Type | Cover_Type == Cover_Type.x)

# Function to calculate the mode
getmode <- function(x) {
   uniqv <- unique(x)
   uniqv[which.max(tabulate(match(x, uniqv)))]
}

# Determine which class is the most common and should be the vote for each row
all.preds.attempt$Vote <- apply(all.preds.attempt[,-1], 1, getmode)

# For some reason the data frame gets grouped, so I ungroup it
all.preds.attempt <- all.preds.attempt %>% ungroup()

# Select all the rows where all three models differ
another.test <- all.preds[!(all.preds$Id %in% all.preds.attempt$Id),]
#Randomly select which model will be used
another.test$random <- sample(1:3, nrow(another.test), replace = TRUE)

# Choose vote based off of the model used and then drop the variable that specifies the value to use
another.test2 <- another.test %>% group_by(Id) %>%
  mutate(Vote = ifelse(random == 1, as.character(Cover_Type.x), ifelse(random == 2, as.character(Cover_Type.y), as.character(Cover_Type)))) %>%
  ungroup() %>%
  select(-random)

# Put back together and arrange by ID
all.preds.together <- all.preds.attempt %>% bind_rows(another.test2) %>% arrange(Id)
all.preds.together <- all.preds.together %>% select(Id, Vote) %>%
  rename(Cover_Type = Vote)

write_csv(all.preds.together, "voting-forest.csv")

```

```{r}

# This is the same as above, but I tried have the better random forest be chosen when two or more didn't agree

all.preds <- gbm %>% left_join(rf, by = "Id") %>% left_join(reg.rf, by = "Id")
all.preds.attempt <- all.preds %>% filter(Cover_Type.x == Cover_Type.y | Cover_Type.y == Cover_Type | Cover_Type == Cover_Type.x)

# Calculate most common and ungroup 
all.preds.attempt$Vote <- apply(all.preds.attempt[,-1], 1, getmode)

all.preds.attempt <- all.preds.attempt %>% ungroup()

# Take rows where all 3 are different and change them to the prediction from the best random forest
another.test <- all.preds[!(all.preds$Id %in% all.preds.attempt$Id),]
another.test2 <- another.test %>% group_by(Id) %>%
  mutate(Vote = Cover_Type.y) %>%
  ungroup()

# Combine back together and create output file
all.preds.together <- all.preds.attempt %>% bind_rows(another.test2) %>% arrange(Id)

all.preds.together <- all.preds.together %>% select(Id, Vote) %>%
  rename(Cover_Type = Vote)

write_csv(all.preds.together, "voting-forest2.csv")
```

